The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-23 manuscript.Rmd
update enterprise section
restructure, shorten a bit

R has been historically viewed as a tool for analysis and scientific research--not for creating software that corporations can rely on to continously run in real time.
However, thanks to advancements in R running as a web service, along with with the ability to deploy R in Docker containers, modern enterprises are now capable of having real-time machine learning powered by R.
One of the first implementations of having R respond to client reqests were TCP/IP and local socket servers, such as \CRANpkg{Rserve} and \CRANpkg{svSocket}, which allowed other local or remote processes to make binary requests to interact with R without the need to link to R at compile time and connected other programming languages with R.
R could also empower web services since at least 2007, when [rApache](http://www.rapache.net) was presented at the useR! conference \citep{horner_rapache_2013}.
<!-- website seems down, https://web.archive.org/web/20190627101203/http://rapache.net/manual.html -->
rApache runs an R interpreter inside the Apache web server, which makes it possible and very efficient to build scalable R-based APIs and web services.

Such rApache applications have been used at [rapporter.net](http://rapporter.net), transforming parametrized R Markdown-like documents into interactive web applications for reporting.
A similar design was implemented at [CARD.com](https://www.card.com/), where the optimization of the acquisition flow and the real-time analysis of debit card transactions both relied on an efficient combination of \CRANpkg{Rserve} and rApache running in Docker containers:
the R session behind Rserve acted as a read-only, extremely fast and scalabale, in-memory database for the many concurrent rApache processes responding to the live-scoring requests of various divisions of the company.
Similar dockerized R scripts were responsible for the ETL processes and even the client-facing email, text message and push notification alerts sent in real-time based on card transactions.
The related Docker images were made available at [https://github.com/cardcorp/card-rocker](https://github.com/cardcorp/card-rocker).
The images extend `rocker/r-base`, e.g., with an SSH client and a workaround for being able to mount SSH keys from the host, Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required by the AWS SDK.
The AWS SDK allowed to run R consumers reading from real-time data processing streams of [AWS Kinesis](https://aws.amazon.com/kinesis/)\footnote{See useR!2017 talk _Stream processing with R in AWS_ at \href{https://static.sched.com/hosted_files/user2017/2f/AWR%20Kinesis%20at%20useR%202017.pdf}{https://static.sched.com/hosted_files/user2017/2f/AWR Kinesis at use 02017.pdf}.}.
The applications were deployed on Amazon Elastic Container Service ([ECS](https://aws.amazon.com/ecs/)).
The main learnings from using R in Docker was the imporance of not only pinning the R package versions via MRAN, but also moving away from Debian testing to an LTS distribution.
For the use case at hand, the priority was to have more control over upstream updates, and to minimize the risk of breaking the automated builds of the Docker images and production jobs.

A more recent related work is the \CRANpkg{plumber} ([https://www.rplumber.io/](https://www.rplumber.io/)) package, which makes it even easier for R to be used to host web services through RESTful APIs.
When R is running on a server, a different server can ask R to make a computation through an HTTP request and get an immediate response.
For instance, if a data scientist at a company creates a machine learning model to tell which customers are likely to make a repeat purchase, if they serve the model through a plumber API then someone else on the marketing team could write software that sends different emails depending on that real-time prediction.
Plumber syntax is simple to read and understand: the API is created by commenting above function names in an R script.

```
#* Return the sum of two numbers
#* @get /random
function(){
  runif()
}
```
_Example of a plumber API to generated a random number by visiting `http://127.0.0.1/random`_

Plumber becomes far more powerful with the addition of Docker.
Docker can be used to create images that [include both R and plumber](https://www.rplumber.io/docs/hosting.html#docker).
Then when the image is deployed, that creates the web endpoints that other services can hit and call the R code.
This pattern of deploying code mimics those used by software engineering services created in Java, Python, and other modern languages.
By using Docker to deploy plumber APIs, R can be used alongside those langages as a first class member of a software engineering technical stack.

At T-Mobile, R is used to power real-time services that are called over a million times a day.
T-Mobile created a set of neural network machine learning natural language processing models to help customer care agents manage text-based messages for customers \citep{t-mobile_enterprise_2018}.
The models are convolutional neural networks that use the RStudio Keras package and are built on top of a Rocker Docker image.
Since the models power tools for agents and customers, they need to have extremely high uptime and reliability--and they found that R is able to perform.
The AI @ T-Mobile team open sourced their Docker images [on GitHub](https://github.com/tmobile/r-tensorflow-api).

[RestRserve](http://restrserve.org) is an alternative R web API framework with the main focus on being high-performant and robust.
It also ships automated Docker builds at https://hub.docker.com/r/rexyai/restrserve/dockerfile.

>>>>

R has been historically viewed as a tool for analysis and scientific research, but not for creating software that corporations can rely on for production services.
However, thanks to advancements in R running as a web service, along with with the ability to deploy R in Docker containers, modern enterprises are now capable of having real-time machine learning powered by R.
A number of packages and projects enabled R to respond to client reqests over TCP/IP and local socket servers, such as \CRANpkg{Rserve}, \CRANpkg{svSocket}, [rApache](http://www.rapache.net) and more recently \CRANpkg{plumber} ([https://www.rplumber.io/](https://www.rplumber.io/)) and \pkg{RestRserve} ([http://restrserve.org](http://restrserve.org)), which even processes incoming requests in parallel with forked processes using \CRANpkg{Rserve}.
The latter two also provide documentation for deployment with Docker or ready to use automated builds of images\footnote{See \href{https://www.rplumber.io/docs/hosting.html#docker}{https://www.rplumber.io/docs/hosting.html#docker}, \href{https://hub.docker.com/r/trestletech/plumber/}{https://hub.docker.com/r/trestletech/plumber/} and \href{https://hub.docker.com/r/rexyai/restrserve/}{https://hub.docker.com/r/rexyai/restrserve/}.}.
These software allow other (remote) processes and programming languages to interact with R and to expose R-based function in a service architecture with HTTP APIs.
APIs based on these package can be deployed with scalability and high availability using containers.
This pattern of deploying code matches those used by software engineering services created in more established languages in the enterprise domain, such as Java or Python, and R can be used alongside those langages as a first class member of a software engineering technical stack.

CARD.com implemented a web application for the optimization of the acquisition flow and the real-time analysis of debit card transactions.
The software used \CRANpkg{Rserve} and rApache and was deployed in Docker containers.
The R session behind \CRANpkg{Rserve} acted as a read-only in-memory database, which was extremely fast and scalabale, for the many concurrent rApache processes responding to the live-scoring requests of various divisions of the company.
Similarly dockerized R scripts were responsible for the ETL processes and even the client-facing email, text message and push notification alerts sent in real-time based on card transactions.
The related Docker images were made available at [https://github.com/cardcorp/card-rocker](https://github.com/cardcorp/card-rocker).
The images extend `rocker/r-base` and additionally entailed an SSH client and a workaround for being able to mount SSH keys from the host, Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required by the AWS SDK.
The AWS SDK allowed to run R consumers reading from real-time data processing streams of [AWS Kinesis](https://aws.amazon.com/kinesis/)\footnote{See useR!2017 talk _Stream processing with R in AWS_ at \href{https://static.sched.com/hosted_files/user2017/2f/AWR%20Kinesis%20at%20useR%202017.pdf}{https://static.sched.com/hosted_files/user2017/2f/AWR Kinesis at use 02017.pdf}.}.
The applications were deployed on Amazon Elastic Container Service ([ECS](https://aws.amazon.com/ecs/)).
The main learnings from using R in Docker was the imporance of not only pinning the R package versions via MRAN, but also moving away from Debian testing to a distribution with long-term support.
For the use case at hand, this switch served the priority to have more control over upstream updates, and to minimize the risk of breaking the automated builds of the Docker images and production jobs.

The AI @ T-Mobile team created a set of neural network machine learning natural language processing models to help customer care agents manage text-based messages for customers \citep{t-mobile_enterprise_2018}.
For example, one model quickly identifies if a message is from a customer or not \citep[see \CRANpkg{Shiny}-based \href{https://secure.message.t-mobile.com/v1/shiny/is-customer/app/}{demo}, ][]{nolis_small_2019}, others tell which customers are likely to make a repeat purchase.
If a data scientist creates a machine learning model and exposes it through a \CRANpkg{plumber} API, then someone else on the marketing team could write software that sends different emails depending on that real-time prediction.
The models are convolutional neural networks that use the \CRANpkg{keras} package and run in a Rocker Docker image.
The corresponding `Dockerfile`s are published [on GitHub](https://github.com/tmobile/r-tensorflow-api).
Since the models power tools for agents and customers, they need to have extremely high uptime and reliability.
The AI @ T-Mobile team found that the models performed well and today these models power real-time services that are called over a million times a day.
